\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-05-09 forwards-and-backwards.tex
edits 2

\begin{definition}[Model parameterization and assumptions]
    \label{def:model_setup}
    Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
    are bounded, uniformly continuous functions
    from $\IR^d \times \IR_{\ge 0}$ to $\IR_{\ge 0}$.
    Furthermore, suppose $r$ takes values in $[0, 1]$
    and has uniformly bounded second derivatives,
    and that $(x, m) \mapsto m^2 \gamma(x, m)$ is unformly bounded.
    Fix nonnegative values $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
    with $\epsilon_r < \epsilon_\gamma$.
    Suppose $F: \IR^d \times \IR_{\ge 0} \to \IR$ is bounded and uniformly continous
    and define
    \begin{align} \label{eqn:mu_defn}
        \mu(x, \eta)
        =
        r(x, \smooth{r} \eta(x))
        \gamma(\smooth{\gamma} \eta(x))
        - \frac{1}{\theta}
        F(x, \smooth{F} \eta(x)) .
    \end{align}
    For each $N > 0$ and $\theta \ge 1/(\epsilon_\gamma - \epsilon_r)$,
    let $\left(\eta^\theta(t)\right)_{t \ge 0}$
    be the continuous-time measure-valued Markov process described above
    with birth rates, establishment probabilities, and death rates
    determined as follows.
    When the state of the process is $\eta \in \measures$,
    offspring of mass $1/N$ are produced at location $x$
    with rate $\theta \gamma(x, \smooth{\gamma} \eta(x)) N \eta_s(dx)$
    and adults die at location $x$ with rate $\theta \mu(x, \smooth{\mu} \eta(x)) N \eta_s(dx)$,
    (i.e., at rates $\theta \gamma$ and $\theta \mu$ per capita respectively).
    Each offspring disperses to a location offset from the parent's location
    by an independent Gaussian with mean $\meanq(x) / \theta$
    and covariance matrix $\covq(x) / \theta$;
    an offspring at $x$ establishes instantaneously
    with probability $r(x, \smooth{r} \eta(x))$, or else dies.
\end{definition}

@@

\comment{I'm not quite sure how we want to frame these two things: probably not as two definitions...}
We will now give another, equivalent definition of $(\eta^{\theta}(t))_{t \geq 0}$,
this time as a solution to a martingale problem.

@@

\comment{
Set-up: there are two dichotomies:
is the limiting process deterministic or a generalized superprocess?
See Figure \ref{fig:super_vs_det_2d} for simulations.
And, are interactions local in the limit or not?
Below we have results for three of these cases;
as for the remaining one (local, superprocess),
it should be possible in 1D, not in higher dimensions (brief note).
}

@@

Theorem \ref{thm:nonlocal_convergence} provided convergence along subsequences;
i.e., for any sequence of values of $\theta$
there is a subsequence that converges.
However, if the limiting process is unique then the convergence is unqualified.
In the deterministic case, we have convergence to a nonlocal PDE.
\comment{Maybe the following two corollaries should be combined?}

\comment{
TODO: insert definition of "weak solution of the PDE \ref{eqn:nonlocal_pde}" heere.
}

\begin{corollary}[Deterministic limits]
    \label{cor:nonlocal_pde_limits}
    \comment{(TODO: remove "has a density" and rephrase appropriately)}
    In the context of Theorem~\ref{thm:nonlocal_convergence},
    if the limit is deterministic (i.e., $\theta / N \to \alpha = 0$), $\eta_0$ has a density,
    and \comment{additional conditions? can we say something like this??}
    then any limit $\eta_t(dx)$ has a density $\varphi(t, x)$,
    which is a weak solution to the nonlocal partial differential equation
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi(t, x)
        =
        r\left(x, \smooth{r} \varphi(t,x) \right)
        \DG^* \left[
            \varphi(t, \cdot)
            \gamma\big( x, \smooth{\gamma} \varphi(t, \cdot) \big)
        \right](x)
        +
        \varphi(t, x)
        F\left(x, \smooth{F} \varphi(t,x) \right)
        .
    \end{equation}
    If equation \eqref{eqn:nonlocal_pde} has a unique solution,
    then $(\eta^\theta_t)_{t \ge 0}$ converges to that solution.

    In particular, if
    \comment{TODO: find some conditions under which this is true?}
    then we have convergence to a unique solution.
\end{corollary}

@@

\comment{TODO: Tom has conditions (coefficient are Lipschitz?) under which
    the solution to \eqref{eqn:nonlocal_pdf} has a density (Thm 2.1 in DDMOT);
    should we insert those here (as a Corollary or Proposition, maybe?)}

>>>>

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are bounded, uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR_{\ge 0}$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
and that $(x, m) \mapsto m^2 \gamma(x, m)$ is unformly bounded.
Fix nonnegative values $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F: \IR^d \times \IR_{\ge 0} \to \IR$ is bounded and uniformly continous
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    r(x, \smooth{r} \eta(x))
    \gamma(\smooth{\gamma} \eta(x))
    - \frac{1}{\theta}
    F(x, \smooth{F} \eta(x)) .
\end{align}
\end{definition}

@@

For each $N > 0$ and $\theta \ge 1/(\epsilon_\gamma - \epsilon_r)$,
the process $\left(\eta^\theta(t)\right)_{t \ge 0}$
is the continuous-time measure-valued Markov process described above
with birth rates, establishment probabilities, and death rates
determined as follows.
When the state of the process is $\eta \in \measures$,
offspring of mass $1/N$ are produced at location $x$
with rate $\theta \gamma(x, \eta) N \eta_s(dx)$
and adults die at location $x$ with rate $\theta \mu(x, \eta) N \eta_s(dx)$,
(i.e., at rates $\theta \gamma$ and $\theta \mu$ per capita respectively).
Each offspring disperses to a location offset from the parent's location
by an independent Gaussian with mean $\meanq(x) / \theta$
and covariance matrix $\covq(x) / \theta$;
an offspring at $x$ establishes instantaneously
with probability $r(x, \eta)$, or else dies.

We have just effectively defined the process, but to be completely clear,
we will now give a formal definition as a solution to a martingale problem.

@@

Our main results depend on two dichotomies:
Is the limiting process deterministic or a (generalized) superprocess?
And, are interactions local in the limit or not?
Below we have results for deterministic limits with local and nonlocal interactions,
and for superprocess limits with nonlocal interactions.
See Figure \ref{fig:super_vs_det_2d} for illustrative simulations.
We suspect that limits are possible the remaining (local, superprocess) case
in only $d=1$, but do not discuss the case further.

@@

Recall when interpreting~\eqref{eqn:mgale_problem}
that for instance $r(x, \eta_s) = r(x, \smooth{r} \eta_s(x))$,
and so $\DG(fr)(x) = \DG(f(\cdot) r(\cdot, \smooth{r} \eta_s(\cdot)))(x)$.

Theorem \ref{thm:nonlocal_convergence} provided convergence along subsequences;
i.e., for any sequence of values of $\theta$
there is a subsequence that converges.
However, if the limiting process is unique then the convergence is unqualified.
In the deterministic case, we have convergence to a nonlocal PDE.

Our results will show that in determinstic limits ``$\eta_t$ solves a PDE'',
but since $\eta_t$ is a measure for which we don't always know there exists a density,
we need to formalize the weak sense in which it solves the PDE.
The notation in the following definition
is written to suggest the case in which $\eta_t(dx)$ has a density $\varphi_t(x)$.

\begin{definition}[Weak solutions]
    \label{defn:weak_solutions}
    We say that $(\eta_t)_{t \ge 0}$, with $\eta_t \in \measures$
    is a \emph{weak solution} to the PDE
    $$
        \dot \varphi = r \DG^*(\gamma \varphi) + \varphi F
    $$
    if for all $f \in C_0^\infty(\IR^d)$,
    $$
        \frac{d}{dt} \langle f, \eta_t \rangle
        =
        \langle
            \gamma \DG(rf) + f F,
            \eta_t
        \rangle .
    $$
    \comment{Write this out with $(x)$'s?}
\end{definition}

\begin{corollary}[Deterministic limits]
    \label{cor:nonlocal_pde_limits}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the limit is deterministic (i.e., $\theta / N \to \alpha = 0$),
    then any limit $\eta_t$ is a weak solution to the nonlocal partial differential equation
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi_t(x)
        =
        r\left(x, \smooth{r} \varphi_t(x) \right)
        \DG^* \left[
            \varphi_t(\cdot)
            \gamma\big( x, \smooth{\gamma} \varphi_t(\cdot) \big)
        \right](x)
        +
        \varphi_t(x)
        F\left(x, \smooth{F} \varphi_t() \right)
        ,
    \end{equation}
    in the sense of Definition~\ref{defn:weak_solutions}.
    If equation \eqref{eqn:nonlocal_pde} has a unique solution,
    then $(\eta^\theta_t)_{t \ge 0}$ converges to that solution.
\end{corollary}

\comment{TODO: find some conditions under which we have convergence to a unique solution?}
\comment{TODO: Tom has conditions (coefficient are Lipschitz?) under which
    the solution to \eqref{eqn:nonlocal_pdf} has a density (Thm 2.1 in DDMOT);
    should we insert those here (as a Corollary or Proposition, maybe?)}

@@
